\papertitle

### 2012-10-13 manuscript.tex
Integrate comments from paper walkthrough

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
\begin{abstract}
The power, size, and weight constraints of micro air vehicles (MAVs) limit 
their on-board sensing and computational resources. Ground vehicles have 
less mobility than MAVs, but relaxed size constraints and typically more 
computing power, creating opportunities for cooperation. In this work, we 
demonstrate cooperative target-seeking between a 13 gram ornithopter MAV 
and a lightweight ground station using computer vision.

We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon a previous design by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
payload of 2.8 grams. We augment the ornithopter's built-in 
gyroscope-based control with a ground station, which has power and weight 
requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
station provides heading estimates to the ornithoper by running a 
real-time motion tracking algorithm on a live video stream. This allows 
the ornithopter to reliably negotiate flight through narrow openings.

We demonstrate the performance of the location estimation and navigation 
algorithms. We model the backwards-reachable region for successfully 
navigating through narrow openings. Finally, we verify this model through 
with experiments with narrow passage traversal.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACM METADATA

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{figure}[!tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird.png}
\caption{H$^2$Bird ornithopter MAV}
\label{fig:h2bird}
\end{figure}
Power, size, and weight constraints for micro air vehicle (MAV) 
platforms are amongst the most restrictive of all robotic platforms. With
significantly limited sensing and computing capabilities, only the most rudimentary
control and state estimation algorithms can be executed on-board these platforms. 
Even more restrictive are the constraints for flapping wing MAVs, 
a class of MAVs that exhibit demonstrate safety and noise profiles superior to those
of rotor craft MAVs, but at the expense of reduced payload weight.

Compounding these engineering restrictions are the many non-linear
flight dynamics exhibited by flapping wing flight~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
Much past research on flapping wing MAVs has been focused on overcoming the 
difficulties associated with controlling and modeling in the face of these 
non-linearities\cite{floreano:flying}. Common approaches to simplify these 
dynamic models include averaging over the wingbeat period~\cite{Schenato:flightcontrol}
\cite{cheng:tandrdamping}, and using linear, low-dimensional models to predict the 
behavior in flight~\cite{Shigeoka:ornithopter}. Advances have lead to the
development of some high-performance platforms, such as the sub-15 gram DelFly 
as described by deCroon et al.~\cite{delfly:design}.

Many of the control laws that previous researches have developed from the
dynamic models of flapping-winged flight, such as periodic forcing and other
control methods as described %TODO you need to use real names, it's not kosher to refer to citation numbers directly
in~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, are computationally complex and
require powerful on-board computing. In our work, we develop a simpler, less
computationally intensive method of control for negotiating narrow passages.

In contrast, there is considerably less previous work on guidance of flapping wing MAVs. 
Baek demonstrated altitude control with an external camera~\cite{baek:altitude}. 
de Croon et al. demonstrated obstacle avoidance with an onboard camera and offboard 
processing~\cite{delfly:avoid}. For sub-20 gram ornithopters, Baek et al. 
\cite{baek:tracking} is one of the few examples of scompletely autonomous target seeking.

There also exists a body of work on guidance and control for fixed wing and rotary wing 
MAV platforms, but many of these methods make extensive use of GPS or motion capture systems
~\cite{kingston:timeattitude}\cite{kanade:3dvision}. While operating indoors, using
GPS to estimate the pose of the MAV is not possible, and motion capture
requires expensive and stationary modification of the environment. The power
and payload limits on flapping wing MAVs make on-board sensing difficult, so multi-agent
cooperative control schemes are apt solutions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robotic and Vision Platforms}
%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper}
% humphrey
To fulfill the need for a lightweight, autonomous flapping wing MAV platform, 
we designed the H$^2$Bird. Built around the Silverlit i-Bird 
RC flyer power train and clap-fling 
wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and can operate for approximately 10 minutes.
%TODO footnote for ImageProc

@@

Rather than power-intensive PC processors, we used the ARM-based 
BeagleBoard single-board computer as reference computation platform. The
BeagleBoard consumes approximately 1 watt while running our computer vision 
and control algorithms. It uses the same processor as popular sub-10 gram 
processor modules---such as the Gumstix Overo and LogicPD Torpedo---which are 
light enough for deployment on millirobotic ground vehicles. We pair the 
BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
quality and resolution to the miniature camera modules available at 
millirobot scale. The BeagleBoard communicates with the H$^2$Bird using a 
USB radio module which implements the IEEE 802.15.4 wireless standard.
%TODO footnotes: Gumstix, LogidPD

Our ground station software is based on the Ubuntu Linux\footnote{\raggedright Ubuntu: 
Canonical Ltd., \href{http://www.ubuntu.com/}{http://www.ubuntu.com/}} distribution
modified with a custom version of the Linux kernel. We use the Python
\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
{http://www.python.org/}} programming language for control, robot communication, 
and telemetry, and the OpenCV\footnote{\raggedright OpenCV: Willow Garage, 
\href{http://opencv.willowgarage.com/wiki/}{http://opencv.willowgarage.com/wiki/}} 
computer vision libraries for image capture and processing.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Concept and Algorithms}
%TODO ryan

@@

Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
high sample rate attitude control. We estimate the vehicle pose by
integrating the IMU, and use a proportional plus integral plus derivative (PID) 
controller for controlling the vehicle.

The high pitch angles flapping wing flyers achieve in routine flight distinguish them 
from other MAVs. Baek demonstrated control on a similar vehicle using PID and an Euler angle
parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of the vehicle orientation 
suffer from singularities around high pitch angles, so we instead 
represent the vehicle orientation with quaternions. An additional benefit of 
the quaternion representation is its ability to represent relative body and 
world coordinate angular displacements compactly with quaternion multiplication.

Our implementation of quaternion-based control is similar to Knoebe's~\cite{knoebe:quatcontrol}:
Given a reference pose represented in quaternion form $q_r$, we define the error as the rotation $q_e$ 
required to reach the reference pose from the estimated pose $q$ (Eq.~\ref{eq:quat_error}). 
\begin{align}
\label{eq:quat_error}
q_e &= q^{\prime}\cdot q_r \\
		&= q_{e,w} + q_{e,x}\hat{i} + q_{e,y}\hat{j} + q_{e,z}\hat{k} \notag 
\end{align}
We then convert this error quaternion to an angle axis representation and use this 
as input to the PID controller (Eqs.~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
These calculations are performed onboard with trigonometric lookup tables for speed.
\begin{align}
\label{eq:quat_linearize}
Q_e &= \alpha\cdot q_e/\sin(\alpha /2) \\
		&= \alpha_{x}\hat{i} + \alpha_{y}\hat{j} + \alpha_{z}\hat{k} \notag
\end{align}
\begin{equation}
\label{eq:quat_angle}
\alpha = 2\arccos(q_{e,w})
\end{equation}

@@

Motion of the H$^2$Bird in the frame is tracked using a particle filter
with an approach similar to that employed by Thrun~\cite{thrun2005probabilistic}. 
Particles are distributed uniformly across the frame on initialization. 
To improve numerical stability, we normalize the weights of all particles on 
each iteration, and only resample the particle population when the mean particle 
weight falls below a pre-determined threshold.

%............................................................................%
\subsubsection{Motion Model}
Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the state of the $i^\text{th}$ particle at time $t$, $\Sigma$ 
is a chosen covariance matrix, and $b$ is vector of the height and width of 
the camera frame.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
The experiments presented in this work use a diagonal covariance matrix 
with a variance of 256, and an $\epsilon$ of 0.3. We find that the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample all of the pixel space. This prevents overconfidence and collapse, 
and decreases reacquisition delay.

%............................................................................%
\subsubsection{Emission Model}
Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
approach to motion tracking via na\"{i}ve background subtraction[CITATION].
\begin{equation}
\label{eq:emission_model}
w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
\end{equation} 
The system provides the emission model with $f_0$, the frame representing 
the background and $f_t$, the most recent frame captured. The model assigns 
a weight $w^{[i]}_t$ to each particle $x^{[i]}_t$, proportional to the color 
distance in RGB space between the pixels of $f_0$ and $f_t$ at the location 
$x^{[i]}_t$. Hence, particles are more likely to be resampled in portions of 
the image which appear different from the background.

For the experiments presented in this work, we assumed that neither the bird 
nor any other moving object were visible in the first frame captured on system 
start. We saved the first frame captured by the system, and provided it as 
the background frame $f_0$ for every iteration of the particle filter. This 
resulted in a sampling approach to na\"{i}ve background subtraction.

Na\"{i}ve background subtraction is known to be brittle to changes in the scene over 
time[CITATION]. However, we found it very efficient and sufficiently reliable for the 
short time spans necessary for H$^2$Bird to successfully negotiate narrow 
openings, which are typically less than five seconds. Substituting $f_0$ 
with a more sophisticated extraction of the background would implement a 
more sophisticated approach to background subtraction, with no changes to 
the underlying motion tracking filter.

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
The H$^2$Bird is remotely guided by commanding relative changes in its
reference orientation. Using relative changes removes the need to estimate the relative
orientation between the ground station and vehicle frames of reference. 
With estimates of the vehicle pose, we calculate an altitude
error and horizontal error. The ground station then commands incremental changes to 
the H$^2$Bird pitch and yaw are according to a PID control law, with positive altitude error
corresponding to negative pitch increments and positive horizontal error
corresponding to negative yaw increments. We window the controller integrator term 
to prevent windup when the bird is being launched, and tuned the derivative term
to dampen oscillations associated with the system latency [?].

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
\caption{Overall block diagram of the cooperative control system.}
\label{fig:block_diagram}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Model}
\label{sec:system_model}

@@

Modeling of the vehicle pose is divided into two independent models for simplicity: one for
yaw motion, and one for translational motion.

We model the yaw motion of the ornithopter with a one-dimensional model to reduce the complexity of the overall model. Matching our physical implementation, we choose the reference heading angle as an input and model the dynamics from the vehicle and on-board PID controller to output the physical heading angle. The model format is shown in Equation~\ref{eq:plain_model}.

@@

To model the translational position of the ornithopter, we use a form of Dubin's car model~\cite{lavalle:planning} with the angular position, $\theta$ as an input rather than a state. The overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the reference heading angle as the input.

\begin{equation}
\label{eq:total_model_plain}
\begin{aligned}
& \dot{x} = sin(\theta)\\
& \dot{y} = cos(\theta)\\
\end{aligned}
\end{equation}

@@

The model of the ground station combines both the camera model and the
computational model for pose estimation. The camera model contains both a
latency, which would occur in between the "camera" and "ornithopter dynamics"
in Fig.~\ref{fig:block_diagram}, and a model for the viewing angle. We model the 
viewing angle for the camera as a pixel map from one edge of a
63$^{\circ}$ viewing triangle to the other. The camera has an image width of
640 pixels. The desired heading angle in degrees for a given distance from the
camera is calculated according to Equation~\ref{eq:desired_heading}, where $d$
is the distance of the robot in front of the camera and $x$ is the horizontal
position of the robot relative to the center of the window in meters. This
relation reflects the fact that the camera cannot determine the depth of the
robot.

\begin{equation} \label{eq:desired_heading} \text{Desired heading} =
\frac{640\text{px}}{2*\text{d}*\tan{\frac{63^{\circ}}{2}}}*\frac{63^{\circ}}{640\text{px}}*\text{x}
\end{equation}

The computational model accounts for the latency added by the operations required to
run the particle filter and calculate the input to send to the robot.

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}

@@

We determined the turning speed and turning radius of the H$^2$Bird
experimentally by measuring the response of the system to the step input of a
clockwise 90 degree turn. To measure the response of the system, we recorded
the angular position data of the ornithopter calculated from the on-board
gyroscope data. We then used
MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} to fit a
simple low-order model to the step response as discussed in Sec.~\ref{sec:model_attitude}.

@@

To measure the capture latency of the camera, we alternatively flashed black
and white color in front of the camera and measured the time lapse before it
detected the color change. We averaged this result over 50 trials to determine
the capture latency.

To compute the particle filter latency, we used system timers to determine the
average time from the start of computation to the end of computation.

%............................................................................%
\subsubsection{Cooperative Visual Servoing}

@@

We tested our cooperative visual servoing system with a simple target seeking
experiment: the H$^2$Bird must fly through a specified obstacle, in this case,
a wooden frame meant to simulate a window. As the H$^2$Bird cannot detect the
target on its own, we used the ground station to provide remote headings for
navigation to the window. A motion capture system in the test room provided
ground truth data. A sketch of the experimental setup is shown in
Fig.~\ref{fig:experiment_cartoon}.

% TODO clarify or roll into paragraph
Each trial included the following steps:
\begin{enumerate}
\item The window is identified by the user. 
\item The H$^2$Bird is released in view of the camera by hand at the desired starting grid point.
\item The H$^2$Bird attempts to fly to the window with remote guidance from
the base station.
\end{enumerate}

In total, we ran 80 trials over a variety of starting positions. The goal of
our experiments was to verify the model described in
Secttion~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a grid
along the edges of the camera viewplane in 0.05 meter increments from the
window to determine the success rate on the edges of the testing space. We
conducted 20 additional trials directly in front of the camera to determine
the success rate near the center of the testing space. After every 5 trials,
we replaced the current battery with a fully charged one.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
\caption{Conceptual sketch of experimental setup.}
\label{fig:experiment_cartoon}
\end{figure}
%TODO label experiment cartoon

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
\begin{abstract}
The power, size, and weight constraints of micro air vehicles (MAVs) limit 
their on-board sensing and computational resources. Ground vehicles have 
less mobility than MAVs, but relaxed size constraints and typically more 
computing power. These specializations present many opportunities for cooperation. In this work, we 
demonstrate cooperative target-seeking between a 13 gram ornithopter MAV 
and a lightweight ground station using computer vision. We develop models for the ornithopter, ground station, and cooperative system dynamics. We determine model parameters of the real systems through experimental system identification. Finally, we verify those models using experiments on narrow passage traversal, and arrive at a cooperative system model which accurately predicts the backwards-reachable region for successfully negotiating ornithopther flight through narrow openings.

We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon a previous design by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
payload of 2.8 grams. We augment the ornithopter's built-in 
gyroscope-based control with a lightweight ground station design, which has power and weight 
requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
station provides heading estimates to the ornithoper by running a 
real-time motion tracking algorithm on a live video stream. This allows 
the ornithopter to reliably negotiate flight through narrow openings. We demonstrate the performance of the location estimation and navigation  algorithms.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACM METADATA

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{figure}[!tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird.png}
\caption{H$^2$Bird ornithopter MAV}
\label{fig:h2bird}
\end{figure}
Power, size, and weight constraints on micro air vehicles (MAV) 
significantly limit sensing their and computational capabilities. They can execute only the most rudimentary
control and state estimation algorithms in real-time. 
The constraints on flapping-wing MAVs are even more restrictive. 
Though they exhibit demonstrate safety and noise profiles superior to those
of rotor craft MAVs, this comes at the expense of reduced payload weight.

The non-linear
flight dynamics exhibited of flapping wing flight compounding constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
Much past research on flapping wing MAVs focused on overcoming the 
difficulties associated with controlling and modeling flapping-wing vehicles in the face of this
non-linearity\cite{floreano:flying}. Common approaches to simplify these 
dynamic models include averaging over the wingbeat period~\cite{Schenato:flightcontrol}
\cite{cheng:tandrdamping}, and using linear, low-dimensional models to predict flight behavior~\cite{Shigeoka:ornithopter}. These advances lead to the
development of some high-performance platforms, such as the sub-15 gram DelFly 
described by deCroon et al.~\cite{delfly:design}.

Many control laws that previous researchers developed from
dynamic models of flapping-winged flight, such as periodic forcing and other
control methods described %TODO you need to use real names, it's not kosher to refer to citation numbers directly
by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
require powerful on-board computing. In this work, we develop a simpler, less
computationally intensive method of control for negotiating narrow passages, by exploiting cooperation between specialized robotic agents.

In contrast to flight control, there is considerably less previous work on guidance and of flapping-wing MAVs. 
Baek demonstrated altitude control with an external camera~\cite{baek:altitude}. 
de Croon et al. demonstrated obstacle avoidance with an onboard camera and offboard 
processing~\cite{delfly:avoid}. For sub-20 gram ornithopters, Baek et al. 
\cite{baek:tracking} provides one of the few examples of completely autonomous target seeking for vehicles at this scale.

There also exists a body of work on the guidance and control of fixed- and rotary-wing 
MAV platforms, but many of these methods make extensive use of GPS or motion capture systems
~\cite{kingston:timeattitude}\cite{kanade:3dvision}. Indoors, using
GPS to estimate the pose of an MAV is impossible, and motion capture
requires extensive and stationary modification of the environment. The power
and payload limits on flapping wing MAVs make complex on-board sensing difficult, so multi-agent
cooperative control schemes are apt solutions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robotic and Vision Platforms}
%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper}
% humphrey
To explore these cooperative control concepts, we designed a new flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{h2bird}). Built around the Silverlit i-Bird 
RC flyer power train and clap-fling 
wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes (Fig.~\ref{h2Bird_axes}. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
controller consists of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and operates for approximately 10 minutes.

@@

Rather than power-intensive PC processors, we used the ARM-based 
BeagleBoard single-board computer as reference computation platform. The
BeagleBoard consumes approximately one watt while running our computer vision 
and control algorithms. It uses the same processor as popular sub-10 gram 
processor modules---such as the Gumstix Overo and LogicPD Torpedo---which are 
light enough for deployment on millirobotic ground vehicles. We pair the 
BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
quality and resolution to the miniature camera modules available at 
millirobot scale. The BeagleBoard communicates with the H$^2$Bird with a
USB radio module which implements the IEEE 802.15.4 wireless standard.
%TODO footnotes: Gumstix, LogidPD

Our ground station software is based on the Ubuntu Linux\footnote{\raggedright Ubuntu: 
Canonical Ltd., \href{http://www.ubuntu.com/}{http://www.ubuntu.com/}} distribution
modified with a custom version of the Linux kernel. We use the Python\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
{http://www.python.org/}} programming language for control, robot communication, 
and telemetry, and the OpenCV\footnote{\raggedright OpenCV: Willow Garage, 
\href{http://opencv.willowgarage.com/wiki/}{http://opencv.willowgarage.com/wiki/}} 
computer vision libraries for image capture and processing.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Concept and Algorithms}
%TODO ryan

@@

We implement attitude estimation and control of the H$^2$Bird on-board to achieve 
high sample-rate attitude control. We estimate the vehicle pose by
integrating the IMU, and use a proportional-integral-derivative (PID) 
controller to feed these estimates back to the attitude control flight surfaces (Figs.\ref{process_flow} and~\ref{block_diagram}).

The high pitch angles flapping-wing MAVs regularly operate at distinguish them 
from other MAVs. Baek demonstrated control on a similar vehicle using PID and an Euler angle
parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of vehicle orientation 
suffer from singularities around high pitch angles. So we instead 
represent the vehicle orientation with quaternions. An additional benefit of 
the quaternion representation is its ability to represent relative body and 
world coordinate angular displacements compactly with quaternion multiplication.

Our implementation of quaternion-based control is similar to Knoebe's~\cite{knoebe:quatcontrol}:
Given a reference pose represented in quaternion form $q_r$, we define the error as the rotation $q_e$ 
required to reach the reference pose from the estimated pose $q$ (Eq.~\ref{eq:quat_error}). 
\begin{align}
\label{eq:quat_error}
q_e &= q^{\prime}\cdot q_r \\
		&= q_{e,w} + q_{e,x}\hat{i} + q_{e,y}\hat{j} + q_{e,z}\hat{k} \notag 
\end{align}
We then convert this error quaternion to an angle axis representation and use it 
as input to the PID controller (Eqs.~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
We perform these calculations on-board, with trigonometric lookup tables for real-time performance.
\begin{align}
\label{eq:quat_linearize}
Q_e &= \alpha\cdot q_e/\sin(\alpha /2) \\
		&= \alpha_{x}\hat{i} + \alpha_{y}\hat{j} + \alpha_{z}\hat{k} \notag
\end{align}
\begin{equation}
\label{eq:quat_angle}
\alpha = 2\arccos(q_{e,w})
\end{equation}

@@

Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a particle filter ~\cite{thrun2005probabilistic}. 
We initialize the particles uniformly across the frame before tracking begins. 
To improve numerical stability, we normalize the weights of all particles on 
each iteration, and only resample the particle population when the mean particle 
weight falls below a pre-determined threshold.

%............................................................................%
\subsubsection{Motion Model}
Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at time $t$, $\Sigma$ 
is a chosen covariance matrix, and $b$ is vector of the bounds of the pixel space.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
In the experiments we present in this work, we used a diagonal covariance matrix 
with a uniform variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample broadly across of the pixel space. This helps prevent overconfidence and collapse, 
and decreases reacquisition delay.

%............................................................................%
\subsubsection{Emission Model}
Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
approach to motion tracking via na\"{i}ve background subtraction, e.g. frame differencing[CITATION].
On each iteration, the system provides the emission model with $f_0$, the frame representing 
the background and $f_t$, the most recent frame captured. The model assigns 
a weight $w^{[i]}_t$ to each particle $x^{[i]}_t$, proportional to the color 
distance in RGB space between the pixels of $f_0$ and $f_t$ at the location 
$x^{[i]}_t$.
\begin{equation}
\label{eq:emission_model}
w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
\end{equation} 
Hence, particles are more likely to be resampled in regions of 
the image which are dissimilar in color from the background.

During experiments presented in this work, we assumed that neither the bird, 
nor any other moving object, were visible in the frame on system 
start. We saved the first frame captured by the system, and provided it as 
the background frame $f_0$ for every iteration of the particle filter. This choice of $f_0$ 
resulted in a sampling approach to na\"{i}ve background subtraction.

Na\"{i}ve background subtraction is known to be brittle to changes in the scene over 
time[CITATION]. However, we find it very efficient and sufficiently reliable for the 
short time spans necessary for H$^2$Bird to successfully negotiate narrow 
openings, which are typically less than five seconds. Substituting $f_0$ 
with a more sophisticated background extraction would retrain the underlying motion tracking features of the filter, while providing it with a more robust way of differentiating subject motion from the background.

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
The ground station guides the H$^2$Bird by commanding changes to its
reference attitude, as governed by its on-board attitude controllers (Figs.\ref{process_flow} and~\ref{block_diagram}). 
With the estimate of 2D vehicle pose from computer vision, we calculate an altitude
error and horizontal error. The ground station then commands a reference attitude to 
the H$^2$Bird pitch and yaw. according to a PID control law. We apply a fixed-lag window the PID controller integrator term 
to prevent wind-up while we launch the ornithopter. We tuned the proportional and derivative terms of this outer controller 
to dampen the overshoot and drift cause by the cooperative system latency.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
\caption{Overall block diagram of the cooperative control system.}
\label{fig:block_diagram}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Model}
\label{sec:system_model}

@@

We divide our model of the vehicle pose into two independent models for simplicity: one for
yaw motion, and one for translational motion.

We approximate the yaw dynamics of the ornithopter with a one-dimensional model to limit complexity and because it matches our physical implementation. Likewise, we choose the reference heading angle as input to the model and model the dynamics of vehicle and on-board PID controller as one system, which outputs the physical heading angle. We show the resulting model form in Equation~\ref{eq:plain_model}.

@@

To model the translational position of the ornithopter, we use a form of Dubin's car model (Eq.~\ref{eq:total_model_plain})~\cite{lavalle:planning}. 
\begin{equation}
\label{eq:total_model_plain}
\begin{aligned}
& \dot{x} = sin(\theta)\\
& \dot{y} = cos(\theta)\\
\end{aligned}
\end{equation}
Our form uses the angular position, $\theta$ as an input rather than a state. We model the overall motion of the system as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (Eq. \ref{eq:plain_model}) with the reference heading angle as the input.

@@

Our model of the ground station comprises models of both the camera and the
pose estimation computation. The camera model contains both a
latency, which is in series with the path between the ``camera'' and ``ornithopter dynamics''
 blocks in Fig.~\ref{fig:block_diagram}, and a model of the camera's viewing angle constraints.

 We model the 
viewing angle of the camera as a pixel map from one edge of a
63$^{\circ}$ viewing triangle to the other. The camera has an image width of
640 pixels. We calculate the desired heading angle in degrees for a given distance from the
camera according to Equation~\ref{eq:desired_heading}, where $d$
is the distance between the ornithopter and the camera, and $x$ is the horizontal
position of the robot relative to the center of the window in meters. This
explicitly fact that the pose estimation algorithm does not estimate the depth of the
robot.
\begin{equation} \label{eq:desired_heading} \text{Desired heading} =
\frac{640\text{px}}{2*\text{d}*\tan{\frac{63^{\circ}}{2}}}*\frac{63^{\circ}}{640\text{px}}*\text{x}
\end{equation}

We model the pose estimation computation by the by the latency of running 
the particle filter, calculating a new reference attitude, and sending to the ornithopter via radio.

@@

We determined the turning speed and turning radius of the H$^2$Bird
experimentally, by measuring the response of the system to a step input of a
clockwise 90$^{\circ}$ turn. We measured the response of the system by recording
the attitude estimates calculated by the ornithopter using its on-board
gyroscope. We then used
MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} to fit a
simple low-order model to the step response, as discussed in Sec.~\ref{sec:model_attitude}.

@@

To measure the capture latency of the camera, we wrote a simple test 
application which changes a large colored rectangle on the ground station's screen, then uses the camera to detect that change. By noting the value of the system timer just before we change the screen, and just after the camera detects that change, the application arrives at an estimate of the video capture pipeline latency. We averaged these result over several hundred rapidly-executed trials to determine the capture latency we used to model the system..

To compute the particle filter and controller latency, we again used system timers to determine the
average time between the start and end of computation.

@@

We tested our cooperative visual servoing system with a simple target seeking
experiment: the H$^2$Bird must fly through a specified obstacle, in this case,
a wooden frame we used to simulate a window. As the H$^2$Bird cannot detect the
target on its own, we used the cooperative control with the ground station, which to provided remote guidance to H$^2$Bird to allow it to navigate to the window.
We used a motion capture system to record 
ground truth data. We present a sketch of this experimental setup in
Fig.~\ref{fig:experiment_cartoon}.

We performed each trial as follows (see Fig.~\ref{fig:pf_screencap}):
\begin{enumerate}
\item The window is identified by the user. 
\item We release H$^2$Bird by hand, in view of the camera at the desired starting grid point.
\item The H$^2$Bird attempts to fly through the window with, with remote guidance from
the ground station.
\end{enumerate}

In total, we ran 80 trials over a variety of starting positions. Our goal for the experiments was to verify the model described in
Secttion~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a grid
along the edges of the camera view plane, and in 0.05 meter increments from the
window, in an effort to determine the success rate on the edges of the testing space. We
conducted 20 additional trials directly in front of the camera to determine
the success rate near the center of viewing plane. After every 5 trials,
we replaced the ornithopter's battery with a fully charged one.
